---
title: "Languages of COVID-19"
author: "Tiago Lubiana"
date: "8/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Wikidata enables multilangual integration. Let's see the usage of different langagues for concepts related to COVID-19. 

First, lets query Wikidata for the counts of languages of 
ITEMS related to either

* The virus,
* The pandemic or 
* The disease

```{r}
# install.packages("WikidataQueryServiceR")
library(WikidataQueryServiceR)
library(ggplot2)
library(dplyr)
library(forcats)

collect <- function(language_colector, df) {
  new = head(df)$lang
  return(unique(c(language_colector, new)))

} 

plot_query <- function(df, y_limit, title) {
  
  mini_df <- head(df)
  mini_df$lang <- as.factor(mini_df$lang)

  mini_df %>%
    mutate(lang = fct_reorder(lang, desc(count))) %>%
    ggplot(aes(x = lang, y = count))+ 
      geom_point() + 
      geom_text(aes(label=count),hjust=0.5, vjust=-1) +
    ylim(0, y_limit) +
    ylab("# uses of each language") + 
    xlab("") +
    theme_classic() + 
    theme(plot.title = element_text(hjust = 0.5),
          axis.title = element_text(size=12),
          axis.text.x = element_text(size=14),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank()) +
    labs(title = title)

}

# Initializing the language colector
language_colector <- c()
```

```{r echo=FALSE}
# Query URL = https://w.wiki/ZdJ

query = '# Core concept as a SUBJECT of the triple
# Language availability of OBJECT

SELECT ?lang (COUNT(*) AS ?count) WHERE {
  {wd:Q84263196 ?prop ?x.} UNION {wd:Q82069695 ?prop ?x.} UNION {wd:Q81068910 ?prop ?x.}
  ?x rdfs:label ?label.
  BIND(LANG(?label) AS ?lang)
  }
GROUP BY ?lang
ORDER BY DESC(?count)
'

items_as_object <- query_wikidata(query)


language_colector <- collect(language_colector, items_as_object)

plot_query(items_as_object, 350, "Language of items (COVID-19 as subject)")
p1 <- plot_query(items_as_object, 350, "Language of items (COVID-19 as subject)")
 
```

Now the same, but with the core concept as OBJECT of the triple. 

```{r}
# Query URL: https://w.wiki/Zd3
# Core concept as a OBJECT of the triple
# Language availability of SUBJECT

query = 'SELECT ?lang (COUNT(*) AS ?count) WHERE {
  {?x ?prop wd:Q84263196.} UNION {?x ?prop wd:Q82069695.} UNION {?x ?prop wd:Q81068910.}
  ?x rdfs:label ?label.
  BIND(LANG(?label) AS ?lang)
  }
GROUP BY ?lang
ORDER BY DESC(?count)
'

items_as_subject <- query_wikidata(query)


language_colector <- collect(language_colector, items_as_subject)
plot_query(items_as_subject, 80000, "Language of items (COVID-19 as object)")
p2 <- plot_query(items_as_subject, 80000, "Language of items (COVID-19 as object)")

```

Now lets look at the language coverage of PROPERTIES. 


```{r}
# Query URL: https://w.wiki/Zd4

query = '# Core concept as a SUBJECT of the triple
# Language availability of PROPERTY

SELECT ?lang (COUNT(*) AS ?count) WHERE {
  {wd:Q84263196 ?prop ?x.} UNION {wd:Q82069695 ?prop ?x.} UNION {wd:Q81068910 ?prop ?x.}
  ?p wikibase:directClaim ?prop.
  ?p rdfs:label ?label.
  BIND(LANG(?label) AS ?lang)
  }
GROUP BY ?lang
ORDER BY DESC(?count)
'

properties_when_core_are_subjects <- query_wikidata(query)

language_colector <- collect(language_colector, properties_when_core_are_subjects)
plot_query(properties_when_core_are_subjects, 800, "Language of properties (COVID-19 as subject)")
p3 <- plot_query(properties_when_core_are_subjects, 800, "Language of properties (COVID-19 as subject)")

```

```{r}
# Query URL: https://w.wiki/Zd6
# This query is too heavy for the API call
# The result (https://w.wiki/ZdA) was manually downloaded
collect <- function(language_colector, df) {
  new = head(df)$lang
  return(unique(c(language_colector, new)))

} 
properties_when_core_are_objects <- read.csv("query_for_properties.csv", stringsAsFactors = FALSE)

language_colector <- collect(language_colector, properties_when_core_are_objects)
plot_query(properties_when_core_are_objects, 80000, "Language of properties (COVID-19 as object)")
p4 <- plot_query(properties_when_core_are_objects, 80000, "Language of properties (COVID-19 as object)")
```
Now let's put all plots together:

```{r}
library(patchwork)
p = (p1 | p2) / (p3 | p4)
p
```


Almost done! Now we need to add a legend for each language in the plot. 




```{r}
# Query URL = https://w.wiki/ZdP



query = '# Language codes
SELECT ?langLabel ?langcode
WHERE
{
  ?lang wdt:P424 ?langcode .
  ?lang wdt:P31/wd:p279* wd:Q34770 .
  SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
}'
  
  language_codes <- query_wikidata(query)
  
  language_codes <-
    language_codes %>% filter(langcode %in% language_colector) %>% arrange(langcode)
  
  
  update_description <- function(id, description, code, language_name, n_of_languages) {
      half_n_of_languages <- ceiling(n_of_languages / 2)
      description <- paste0(description, code, ": ", language_name)
    
      if (id != n_of_languages & id != half_n_of_languages) {
        description <- paste0(description," | ")
      } 
      
      if (id == half_n_of_languages) {
         description <- paste0(description," \n")
      } 
      return(description)

    
    }
  
  n_of_languages <- nrow(language_codes)

  description = ""
  for (id in 1:n_of_languages) {
    code = language_codes$langcode[id]
    language_name = language_codes$langLabel[id]
    print(id)
    description = update_description(id, description, code, language_name, n_of_languages)

  }
  print(description)
  p = p + plot_annotation(caption = description) & 
  theme(plot.caption = element_text(size = 12))

  ggsave("use_per_language.png", p,  width = 10, height = 6)

```



